\thetitle

### 2020-07-24 writing/introduction.tex
adding to the intro academic claim about useless blame

From the very beginning, the theory of gradual typing has come with a blame
theorem~\cite{mf-toplas-2009, tf-dls-2006}. ``Well-typed [components]\footnote{The
original authors got this word wrong.} can't be blamed'' turned the theorem into a
slogan~\cite{wf-esop-2009}. Indeed, academic implementations---for example,
Reticulated Python~\cite{vsc-dls-2019, vss-popl-2017, vksb-dls-2014} and Typed
Racket~\cite{tf-dls-2006,tf-popl-2008,tfffgksst-snapl-2017,tf-icfp-2010}---come
with sophisticated blame assignment strategies. Theoreticians firmly embrace the
idea that blame can help practicing programmers find impedance mismatches between
the types they added to a software system and the behavior of the remaining untyped
components.

Practical gradual type systems ignore blame assignment almost completely.
Industrial systems such as Flow,\footnote{\url{https://flow.org}}
Hack,\footnote{\url{https://hacklang.org}} or
TypeScript\footnote{\url{https://www.typescriptlang.org}} exploit types for IDE actions and for finding typos in
code. Then these compilers remove types and rely on the built-in safety checks of
the underlying programming language to catch any problems.

The contrast between academic and industrial systems raises the question 
\begin{quote}
 \it
 whether blame assignment adds any value to a gradually typed language, \\
 especially for the benefit of the working programmer.
\end{quote}
Given the prominence of blame in academia and its complete absence in
industrial systems, it comes as an even bigger surprise that the research
literature and the industrial blog world do not present or discuss any
possible answers.  Instead, when language designers decide to look into the
question, they seem to answer it one way or another without any scientific
justification. As a matter of fact, the research community simply does not have a
method for evaluating blame, and given the question's practical
relevance, this a significant problem.

>>>>

From the very beginning, the theory of gradual typing has come with a
blame theorem~\cite{mf-toplas-2009, tf-dls-2006}. ``Well-typed
[components]\footnote{The original authors got this word wrong.} can't be
blamed'' turned the theorem into a slogan~\cite{wf-esop-2009}. Indeed,
academic implementations---for example, Reticulated
Python~\cite{vsc-dls-2019, vss-popl-2017, vksb-dls-2014} and Typed
Racket~\cite{tf-dls-2006,tf-popl-2008,tfffgksst-snapl-2017,tf-icfp-2010}---come
with sophisticated blame assignment strategies. The academics that
created them firmly embrace the idea that blame can help practicing
programmers find impedance mismatches between the types they added to a
software system and the behavior of the remaining untyped components.

However, many academics view blame as an interesting but laborious
orthogonal add-on to 
gradual typing. Thus blame is often omitted from theoretical models and
academic implementations alike.
Other academics go one step further and doubt its pragmatic significance altogether~\cite{cc-snapl-19}.

In line with the latter, practical gradual type systems ignore blame assignment almost completely.
Industrial systems such as Flow,\footnote{\url{https://flow.org}}
Hack,\footnote{\url{https://hacklang.org}} or
TypeScript\footnote{\url{https://www.typescriptlang.org}} exploit types for IDE actions and for finding typos in
code. Then these compilers remove types and rely on the built-in safety checks of
the underlying programming language to catch any problems.

The divisions in the academic and industrial world of gradual typing raise the question 
\begin{quote}
 \it
 whether blame assignment adds any value to a gradually typed language, \\
 especially for the benefit of the working programmer.
\end{quote}
Given the prominence of blame in academia and its complete absence in
industrial systems, it comes as an even bigger surprise that the research
literature and the industrial blog world do not present or discuss any
possible answers.  Instead, when language designers decide to look into the
question, they seem to answer it one way or another without any scientific
justification. As a matter of fact, the research community simply does not have a
method for evaluating blame, and given the question's practical
relevance, this a significant problem.
